---
title: "corss-compare"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r, message = F, echo = F, warning = F}
library(ProbBayes)
library(tidyverse)
library(bayesrules)
library(patchwork)
library(ggplot2)
library(runjags)
library(coda)      
library(bayesplot) 
```


# data wrangling 
```{r}
between<- read.csv("between.csv")
#View(between)
cross<- between %>% mutate(Test_Rate = Total_Tests/Enrollment)
cross<- cross %>% mutate(type_num = ifelse(Type  == "LAC", 1, 0))

```

# JAGS
```{r}
modelString <-"
model {
## likelihood
for (i in 1:N){
   y[i] ~ dbin(p[i], n[i])
}

## priors and regression
for (i in 1:N){
   p[i] ~ dbeta(a[i], b[i])
   a[i] <- phi[i] * mu[i]
   b[i] <- phi[i] * (1 - mu[i])
   logit(mu[i]) <- beta0 +beta1*type[i] +beta2*pop[i] +beta3*test_rate[i]
   phi[i] <- exp(logeta[i])
   logeta[i] ~ dlogis(logn, 1)
   
}
## hyperpriors

beta0 ~ dnorm(0, 0.01)
beta1 ~ dnorm(0, 0.001)
beta2 ~ dnorm(0, 0.001)
beta3 ~ dnorm(0, 0.001)
}
"

y <- cross$Positive_Results
n <- cross$Total_Tests
type <- cross$type_num
pop <- cross$Enrollment
test_rate<- cross$Test_Rate
N <- length(y)
the_data <- list("y" = y, "n" = n, "N" = N, "type" = type, "pop" = pop, "test_rate" = test_rate,
                 "logn" = log(100))



init = list(
            list(.RNG.name = "base::Wichmann-Hill", .RNG.seed = 987654), 
            list(.RNG.name = "base::Wichmann-Hill", .RNG.seed = 987653),
            list(.RNG.name = "base::Wichmann-Hill", .RNG.seed = 987652)
            )



posterior_school <- run.jags(modelString,
                 data = the_data,
                 n.chains = 3,
                 monitor = c("p"),
                 adapt = 1000,
                 burnin = 5000,
                 sample = 5000,
                 inits = init, silent.jags = TRUE)



```

# diagnostics
```{r}
# p all converged
mcmc_trace(posterior_school$mcmc)
# acf all looks good
mcmc_acf(posterior_school$mcmc)
#looks good

mcmc_dens_overlay(posterior_school$mcmc)

summary(posterior_school$mcmc)
```

# inference
```{r, eval = F, echo = F }
post_intervals <- mcmc_intervals_data(posterior_school$mcmc, regex_pars = "p", prob_outer = 0.9)
slice(post_intervals, 1:30) %>%
  ggplot(
    aes(x = reorder(parameter, (m)), y = (m), ymin = (ll), ymax = (hh))) +
  geom_pointrange() +
  theme_light() +
  xaxis_text(angle = 0, size = 6) +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank()
)+ xlab("School") + ylab("Rate")
```

# betas
```{r}
posterior_beta <- run.jags(modelString,
                 data = the_data,
                 n.chains = 3,
                 monitor = c("beta0", "beta1", "beta2", "beta3"),
                 adapt = 1000,
                 burnin = 5000,
                 sample = 5000,
                 thin = 5,
                 inits = init, silent.jags = TRUE)


# good enough
mcmc_trace(posterior_beta$mcmc)
# acf all looks good
mcmc_acf(posterior_beta$mcmc)
#looks good

mcmc_dens_overlay(posterior_beta$mcmc)
```

# notes
1. the mean of the p for each institution is a regression of the type of institution, the enrollment number, the test rate
2. however, the value of beta2, ie the coefficient of enrollment is too small


```{r}
summary(posterior_beta)
```


# dropped enrollment as a regressor 
```{r}
modelString2 <-"
model {
## likelihood
for (i in 1:N){
   y[i] ~ dbin(p[i], n[i])
}

## priors and regression
for (i in 1:N){
   p[i] ~ dbeta(a[i], b[i])
   a[i] <- phi[i] * mu[i]
   b[i] <- phi[i] * (1 - mu[i])
   logit(mu[i]) <- beta0 +beta1*type[i] +beta3*test_rate[i]
   phi[i] <- exp(logeta[i])
   logeta[i] ~ dlogis(logn, 1)
   
}
## hyperpriors

beta0 ~ dnorm(0, 0.01)
beta1 ~ dnorm(0, 0.001)
beta3 ~ dnorm(0, 0.001)
}
"

y <- cross$Positive_Results
n <- cross$Total_Tests
type <- cross$type_num
pop <- cross$Enrollment
test_rate<- cross$Test_Rate
N <- length(y)
the_data2 <- list("y" = y, "n" = n, "N" = N, "type" = type, "test_rate" = test_rate,
                 "logn" = log(100))



init = list(
            list(.RNG.name = "base::Wichmann-Hill", .RNG.seed = 987654), 
            list(.RNG.name = "base::Wichmann-Hill", .RNG.seed = 987653),
            list(.RNG.name = "base::Wichmann-Hill", .RNG.seed = 987652)
            )



posterior_beta2 <- run.jags(modelString2,
                 data = the_data2,
                 n.chains = 3,
                 monitor = c("beta0", "beta1", "beta3"),
                 adapt = 1000,
                 burnin = 5000,
                 sample = 5000,
                 thin = 5,
                 inits = init, silent.jags = TRUE)

summary(posterior_beta2)


```


```{r}
# good enough
mcmc_trace(posterior_beta2$mcmc)
# acf all looks good -- thin = 5
mcmc_acf(posterior_beta2$mcmc)
#looks good
mcmc_dens_overlay(posterior_beta2$mcmc)
```

```{r}
summary(posterior_beta2)
```



## posterior check

### plot regressions lines
```{r}
#plot 100 posterior model lines for beta3 (beta1 for type is dummy)
posterior_lines_beta3 <- posterior_beta2$mcmc[[1]][1:100,c("beta0", "beta3")] %>%
as.data.frame() %>%
mutate(x = -100, xend = 100, y = beta0 + beta3 * x , yend = beta0 + beta3 * xend)
ggplot(posterior_lines_beta3) +
geom_segment(aes(x = x, y = y, xend = xend, yend = yend), alpha = 0.5)

```

### posterier predictive check
```{r}

modelString3 <-"
model {
## likelihood
for (i in 1:N){
   y[i] ~ dbin(p[i], n[i])
}

## priors and regression
for (i in 1:N){
   p[i] ~ dbeta(a[i], b[i])
   a[i] <- phi[i] * mu[i]
   b[i] <- phi[i] * (1 - mu[i])
   logit(mu[i]) <- beta0 +beta1*type[i] +beta3*test_rate[i]
   phi[i] <- exp(logeta[i])
   logeta[i] ~ dlogis(logn, 1)
   
}
## hyperpriors

beta0 ~ dnorm(0, 0.01)
beta1 ~ dnorm(0, 0.001)
beta3 ~ dnorm(0, 0.001)


#predictive 
for (i in 1:N) {
   y_pred[i] ~ dbin(p[i], n[i])
}

}
"


y <- cross$Positive_Results
n <- cross$Total_Tests
type <- cross$type_num
pop <- cross$Enrollment
test_rate<- cross$Test_Rate
N <- length(y)
the_data2 <- list("y" = y, "n" = n, "N" = N, "type" = type, "test_rate" = test_rate,
                 "logn" = log(100))



init = list(
            list(.RNG.name = "base::Wichmann-Hill", .RNG.seed = 987654), 
            list(.RNG.name = "base::Wichmann-Hill", .RNG.seed = 987653),
            list(.RNG.name = "base::Wichmann-Hill", .RNG.seed = 987652)
            )



posterior_pred <- run.jags(modelString3,
                 data = the_data2,
                 n.chains = 3,
                 monitor = c("y_pred"),
                 adapt = 1000,
                 burnin = 5000,
                 sample = 5000,
                 inits = init, silent.jags = TRUE)

```


```{r}
#??
ppc_dens_overlay(cross$Positive_Results, posterior_pred$mcmc[[1]])
```

```{r}
#actually pretty good
ppc_stat_grouped(y = cross$Positive_Results, 
                 yrep= posterior_pred$mcmc[[1]],
                 group = cross$Institution)
```

## try log and z-score enrollment

```{r}
modelString4 <-"
model {
## likelihood
for (i in 1:N){
   y[i] ~ dbin(p[i], n[i])
}

## priors and regression
for (i in 1:N){
   p[i] ~ dbeta(a[i], b[i])
   a[i] <- phi[i] * mu[i]
   b[i] <- phi[i] * (1 - mu[i])
   logit(mu[i]) <- beta0 +beta1*type[i] +beta2*log_enrol[i]+beta3*test_rate[i]
   phi[i] <- exp(logeta[i])
   logeta[i] ~ dlogis(logn, 1)
   
}
## hyperpriors

beta0 ~ dnorm(0, 0.01)
beta1 ~ dnorm(0, 0.001)
beta2 ~ dnorm(0, 0.001)
beta3 ~ dnorm(0, 0.001)


#predictive 
for (i in 1:N) {
   y_pred[i] ~ dbin(p[i], n[i])
}

}
"


y <- cross$Positive_Results
n <- cross$Total_Tests
type <- cross$type_num
pop <- cross$Enrollment
test_rate<- cross$Test_Rate
log_enrol <- log(cross$Enrollment)
N <- length(y)
the_data4 <- list("y" = y, "n" = n, "N" = N, "type" = type, "test_rate" = test_rate,"log_enrol" = log_enrol,
                 "logn" = log(100))



init = list(
            list(.RNG.name = "base::Wichmann-Hill", .RNG.seed = 987654), 
            list(.RNG.name = "base::Wichmann-Hill", .RNG.seed = 987653),
            list(.RNG.name = "base::Wichmann-Hill", .RNG.seed = 987652)
            )



posterior4 <- run.jags(modelString4,
                 data = the_data4,
                 n.chains = 3,
                 monitor = c("beta0", "beta1", "beta2","beta3"),
                 adapt = 1000,
                 burnin = 5000,
                 sample = 5000,
                 thin = 5,
                 inits = init, silent.jags = TRUE)


```

```{r}
# not good except for beta3
mcmc_trace(posterior4$mcmc)
# acf not good even with thin =5
mcmc_acf(posterior4$mcmc)
#nono
mcmc_dens_overlay(posterior4$mcmc)
```

## z-score enrollment
```{r}
modelString5 <-"
model {
## likelihood
for (i in 1:N){
   y[i] ~ dbin(p[i], n[i])
}

## priors and regression
for (i in 1:N){
   p[i] ~ dbeta(a[i], b[i])
   a[i] <- phi[i] * mu[i]
   b[i] <- phi[i] * (1 - mu[i])
   logit(mu[i]) <- beta0 +beta1*type[i] +beta2*z_enrol[i]+beta3*test_rate[i]
   phi[i] <- exp(logeta[i])
   logeta[i] ~ dlogis(logn, 1)
   
}
## hyperpriors

beta0 ~ dnorm(0, 0.01)
beta1 ~ dnorm(0, 0.001)
beta2 ~ dnorm(0, 0.001)
beta3 ~ dnorm(0, 0.001)


#predictive 
for (i in 1:N) {
   y_pred[i] ~ dbin(p[i], n[i])
}

}
"


y <- cross$Positive_Results
n <- cross$Total_Tests
type <- cross$type_num
pop <- cross$Enrollment
test_rate<- cross$Test_Rate
z_enrol <- (cross$Enrollment -mean(cross$Enrollment))/sqrt(var(cross$Enrollment))
N <- length(y)
the_data5 <- list("y" = y, "n" = n, "N" = N, "type" = type, "test_rate" = test_rate,"z_enrol" = z_enrol,
                 "logn" = log(100))



init = list(
            list(.RNG.name = "base::Wichmann-Hill", .RNG.seed = 987654), 
            list(.RNG.name = "base::Wichmann-Hill", .RNG.seed = 987653),
            list(.RNG.name = "base::Wichmann-Hill", .RNG.seed = 987652)
            )



posterior5 <- run.jags(modelString5,
                 data = the_data5,
                 n.chains = 3,
                 monitor = c("beta0", "beta1", "beta2","beta3"),
                 adapt = 1000,
                 burnin = 5000,
                 sample = 5000,
                 thin = 5,
                 inits = init, silent.jags = TRUE)

summary(posterior5)
```


```{r}
# not good except for beta3
mcmc_trace(posterior5$mcmc)
# acf not good even with thin =5
mcmc_acf(posterior5$mcmc)
#nono
mcmc_dens_overlay(posterior5$mcmc)
```

## diagnose
```{r}
posterior5_2 <- run.jags(modelString5,
                 data = the_data5,
                 n.chains = 3,
                 monitor = c("p"),
                 adapt = 1000,
                 burnin = 5000,
                 sample = 5000,
                 thin = 5,
                 inits = init, silent.jags = TRUE)

# good
mcmc_trace(posterior5_2$mcmc)
# good
mcmc_acf(posterior5_2$mcmc)
#good
mcmc_dens_overlay(posterior5_2$mcmc)



```


```{r}
post_intervals2 <- mcmc_intervals_data(posterior5_2$mcmc, regex_pars = "p", prob_outer = 0.9)
slice(post_intervals2, 1:30) %>%
  ggplot(
    aes(x = reorder(parameter, (m)), y = (m), ymin = (ll), ymax = (hh))) +
  geom_pointrange() +
  theme_light() +
  xaxis_text(angle = 0, size = 6) +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank()
)+ xlab("School") + ylab("Rate")
```

## posterior predictive
```{r}
posterior_pred2 <- run.jags(modelString5,
                 data = the_data5,
                 n.chains = 3,
                 monitor = c("y_pred"),
                 adapt = 1000,
                 burnin = 5000,
                 sample = 5000,
                 thin = 5,
                 inits = init, silent.jags = TRUE)

```

```{r}
# good
ppc_dens_overlay(cross$Positive_Results, posterior_pred2$mcmc[[1]])
```

```{r}
#pretty good
ppc_stat_grouped(y = cross$Positive_Results, 
                 yrep= posterior_pred2$mcmc[[1]],
                 group = cross$Institution)
```

## try just intercept and test_rate
```{r}
modelString5_rate <-"
model {
## likelihood
for (i in 1:N){
   y[i] ~ dbin(p[i], n[i])
}

## priors and regression
for (i in 1:N){
   p[i] ~ dbeta(a[i], b[i])
   a[i] <- phi[i] * mu[i]
   b[i] <- phi[i] * (1 - mu[i])
   logit(mu[i]) <- beta0 +beta3*test_rate[i]
   phi[i] <- exp(logeta[i])
   logeta[i] ~ dlogis(logn, 1)
   
}
## hyperpriors

beta0 ~ dnorm(0, 0.01)
beta3 ~ dnorm(0, 0.001)


#predictive 
for (i in 1:N) {
   y_pred[i] ~ dbin(p[i], n[i])
}

}
"


y <- cross$Positive_Results
n <- cross$Total_Tests
type <- cross$type_num
pop <- cross$Enrollment
test_rate<- cross$Test_Rate
z_enrol <- (cross$Enrollment -mean(cross$Enrollment))/sqrt(var(cross$Enrollment))
N <- length(y)
the_data5_rate <- list("y" = y, "n" = n, "N" = N, "test_rate" = test_rate,
                 "logn" = log(100))



init = list(
            list(.RNG.name = "base::Wichmann-Hill", .RNG.seed = 987654), 
            list(.RNG.name = "base::Wichmann-Hill", .RNG.seed = 987653),
            list(.RNG.name = "base::Wichmann-Hill", .RNG.seed = 987652)
            )



posterior5_rate <- run.jags(modelString5_rate,
                 data = the_data5_rate,
                 n.chains = 3,
                 monitor = c("beta0","beta3"),
                 adapt = 1000,
                 burnin = 5000,
                 sample = 5000,
                 thin = 5,
                 inits = init, silent.jags = TRUE)

summary(posterior5_rate)
```



```{r}
mcmc_trace(posterior5_rate$mcmc)
mcmc_acf(posterior5_rate$mcmc)
mcmc_dens_overlay(posterior5_rate$mcmc)


mcmc_trace(posterior_beta2$mcmc)
# acf all looks good -- thin = 5
mcmc_acf(posterior_beta2$mcmc)
#looks good
mcmc_dens_overlay(posterior_beta2$mcmc)

```





# just LAC

```{r}
lac<- cross %>% filter(cross$Type == "LAC")
```


## same formula for posterior5
```{r}
modelString6 <-"
model {
## likelihood
for (i in 1:N){
   y[i] ~ dbin(p[i], n[i])
}

## priors and regression
for (i in 1:N){
   p[i] ~ dbeta(a[i], b[i])
   a[i] <- phi[i] * mu[i]
   b[i] <- phi[i] * (1 - mu[i])
   logit(mu[i]) <- beta0 +beta2*z_enrol[i]+beta3*test_rate[i]
   phi[i] <- exp(logeta[i])
   logeta[i] ~ dlogis(logn, 1)
   
}
## hyperpriors

beta0 ~ dnorm(0, 0.01)
beta1 ~ dnorm(0, 0.001)
beta2 ~ dnorm(0, 0.001)
beta3 ~ dnorm(0, 0.001)


#predictive 
for (i in 1:N) {
   y_pred[i] ~ dbin(p[i], n[i])
}

}
"


y <- lac$Positive_Results
n <- lac$Total_Tests
pop <- lac$Enrollment
test_rate<- lac$Test_Rate
z_enrol <- (lac$Enrollment -mean(lac$Enrollment))/sqrt(var(lac$Enrollment))
N <- length(y)
the_data6 <- list("y" = y, "n" = n, "N" = N, "test_rate" = test_rate,"z_enrol" = z_enrol,
                 "logn" = log(100))



init = list(
            list(.RNG.name = "base::Wichmann-Hill", .RNG.seed = 987654), 
            list(.RNG.name = "base::Wichmann-Hill", .RNG.seed = 987653),
            list(.RNG.name = "base::Wichmann-Hill", .RNG.seed = 987652)
            )



posterior6 <- run.jags(modelString6,
                 data = the_data6,
                 n.chains = 3,
                 monitor = c("beta0", "beta2","beta3"),
                 adapt = 1000,
                 burnin = 5000,
                 sample = 5000,
                 thin = 5,
                 inits = init, silent.jags = TRUE)

summary(posterior6)

```

## drop test_rate
```{r}
modelString7 <-"
model {
## likelihood
for (i in 1:N){
   y[i] ~ dbin(p[i], n[i])
}

## priors and regression
for (i in 1:N){
   p[i] ~ dbeta(a[i], b[i])
   a[i] <- phi[i] * mu[i]
   b[i] <- phi[i] * (1 - mu[i])
   logit(mu[i]) <- beta0 +beta2*z_enrol[i]
   phi[i] <- exp(logeta[i])
   logeta[i] ~ dlogis(logn, 1)
   
}
## hyperpriors

beta0 ~ dnorm(0, 0.01)
beta2 ~ dnorm(0, 0.001)



#predictive 
for (i in 1:N) {
   y_pred[i] ~ dbin(p[i], n[i])
}

}
"


y <- lac$Positive_Results
n <- lac$Total_Tests
z_enrol <- (lac$Enrollment -mean(lac$Enrollment))/sqrt(var(lac$Enrollment))
N <- length(y)
the_data7 <- list("y" = y, "n" = n, "N" = N, "z_enrol" = z_enrol,
                 "logn" = log(100))



init = list(
            list(.RNG.name = "base::Wichmann-Hill", .RNG.seed = 987654), 
            list(.RNG.name = "base::Wichmann-Hill", .RNG.seed = 987653),
            list(.RNG.name = "base::Wichmann-Hill", .RNG.seed = 987652)
            )



posterior7 <- run.jags(modelString7,
                 data = the_data7,
                 n.chains = 3,
                 monitor = c("beta0", "beta2"),
                 adapt = 1000,
                 burnin = 5000,
                 sample = 5000,
                 thin = 5,
                 inits = init, silent.jags = TRUE)

summary(posterior7)
```


## just test_rate intercept


```{r}
modelString8 <-"
model {
## likelihood
for (i in 1:N){
   y[i] ~ dbin(p[i], n[i])
}

## priors and regression
for (i in 1:N){
   p[i] ~ dbeta(a[i], b[i])
   a[i] <- phi[i] * mu[i]
   b[i] <- phi[i] * (1 - mu[i])
   logit(mu[i]) <- beta0 +beta3*test_rate[i]
   phi[i] <- exp(logeta[i])
   logeta[i] ~ dlogis(logn, 1)
   
}
## hyperpriors

beta0 ~ dnorm(0, 0.01)
beta3 ~ dnorm(0, 0.001)



#predictive 
for (i in 1:N) {
   y_pred[i] ~ dbin(p[i], n[i])
}

}
"


y <- lac$Positive_Results
n <- lac$Total_Tests
test_rate<- lac$Test_Rate
N <- length(y)
the_data8 <- list("y" = y, "n" = n, "N" = N, "test_rate" = test_rate,
                 "logn" = log(100))



init = list(
            list(.RNG.name = "base::Wichmann-Hill", .RNG.seed = 987654), 
            list(.RNG.name = "base::Wichmann-Hill", .RNG.seed = 987653),
            list(.RNG.name = "base::Wichmann-Hill", .RNG.seed = 987652)
            )



posterior8 <- run.jags(modelString8,
                 data = the_data8,
                 n.chains = 3,
                 monitor = c("beta0", "beta3"),
                 adapt = 1000,
                 burnin = 5000,
                 sample = 5000,
                 thin = 5,
                 inits = init, silent.jags = TRUE)

summary(posterior8)
```


# no regressor is fine

```{r}
modelString9 <-"
model {
## likelihood
for (i in 1:N){
   y[i] ~ dbin(p[i], n[i])
}
## priors
for (i in 1:N){
   p[i] ~ dbeta(a, b)
}
## hyperpriors
a <- mu*eta
b <- (1-mu)*eta
mu ~ dbeta(mua, mub)
eta <- exp(logeta)
logeta ~ dlogis(logn, 1)
}
"
y <- lac$Positive_Results
n <- lac$Total_Tests
N <- length(y)
the_data9 <- list("y" = y, "n" = n, "N" = N,
                 "mua" = 1, "mub" = 1,
                 "logn" = log(100))


init = list(
            list(.RNG.name = "base::Wichmann-Hill", .RNG.seed = 987654), 
            list(.RNG.name = "base::Wichmann-Hill", .RNG.seed = 987653),
            list(.RNG.name = "base::Wichmann-Hill", .RNG.seed = 987652)
            )



posterior9 <- run.jags(modelString9,
                 data = the_data9,
                 n.chains = 3,
                 monitor = c("p"),
                 adapt = 1000,
                 burnin = 5000,
                 sample = 5000,
                 inits = init, silent.jags = TRUE)


```

